Optimal (Multiway) Spatial Joins

### 2024-05-25 msj.tex
yufei

%We will find result tuples of each type separately. Before proceeding, however, let us observe a sense of symmetry for Type 2. As mentioned, for each result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both. As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc. It is possible for the result tuple to belong to multiple sub-types simultaneously.

\vgap 
%
% Our subsequent discussion will proceed as follows.  Section~\ref{sec:hv:type1} will explain how to produce all the result tuples of Type 1. Section~\ref{sec:hv:type2} will explain how to produce the result tuples $(r_1, ..., r_{k-2}, h, v)$ of a particular sub-type of Type 2: those tuples where $r_1$ covers the left endpoint of $h$. Other sub-types can be computed using the same algorithm, by symmetry.
%
% \vgap
%
% A remark is in order about duplicate removal. A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types. Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm. However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times. Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to. To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting. One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.

@@

        \item Consider any $k$-tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$. Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$. Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$. Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
    }
\end{lemma}

@@

\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}. First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time. Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates. For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time. Abort the scan as soon as $h \notin \dcross_H(\bm{t})$. This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t}|))$ time. Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.

@@

>>>>

@@

        \item Consider any tuple $\bm{t} \in \J(R'_1, ..., R'_{k-2})$. Let $r' = \gbot(\bm{t})$ and $r^*$ be the top-sliced rectangle of $r'$. Then, we have $\dcross_H(\bm{t})$ $\subseteq \cross_H(r^*)$. Furthermore, if the (horizontal) segments of $\cross_H(r^*)$ are sorted in ascending order of y-coordinate, then $\dcross_H(\bm{t})$ includes a prefix of the sorted order.
    }
\end{lemma}

@@

\extraspacing {\bf Computing the ``d-cross'' Sets.} We are ready to compute $\dcross_H(\bm{t})$, defined in \eqref{eqn:dcross-H}, for any $\bm{t} \in \J(R'_1, ..., R'_{k-2})$, thanks to Statement (2) of Lemma~\ref{lmm:hv:type1:cross-r*}. First, compute $B_\bm{t}$, obtain the rectangle $r' = \gbot(\bm{t})$, and fetch the (already computed) top-sliced rectangle $r^*$ of $r'$; these steps require $O(k)$ time. Then, scan the segments in $\cross_H(r^*)$ in ascending order of their y-coordinates. For each segment $h$ scanned, check whether $h$ belongs to $\dcross_H(\bm{t})$, namely, whether $h$ crosses $B_\bm{t}$ (the reader can verify that $h$ must cross $\gbot(\bm{t})$); this can be done in constant time. Abort the scan as soon as $h \notin \dcross_H(\bm{t})$. This way, we produce $\dcross_H(\bm{t})$ in $O(k + |\dcross_H(\bm{t})|)$ time. Doing so for all $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ takes $O(k \cdot |\J| + \sum_\bm{t} |\dcross_H(t)|) = O(k \cdot \out)$ time, where the derivation used statement (2) of Lemma~\ref{lmm:hv:type1:properties}.

@@

Still, denote by $R_1, ..., R_{k-2}, H$, and $V$ the input sets of the H-V $k$-SJ problem. This section will explain how to find the result tuples of Type 2 as defined in Section~\ref{sec:hv}.

As mentioned before, for a $(k-2)$-tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both. As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc. It is possible for the result tuple to belong to multiple sub-types simultaneously.

Next, we will focus on producing the result tuples of a particular sub-type: those tuples $(r_1, ..., r_{k-2}, h, v)$ where $r_1$ covers the left endpoint of $h$ --- let us denote this set of result tuples as $\J_2$. Other sub-types can be computed using the same algorithm, by symmetry.

A remark is in order about duplicate removal. By finding each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm. However, this does not mean that the tuple needs to be {\em reported} multiple times. Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to. To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting. One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.

\extraspacing {\bf Set $\bm{H'}$.}
